---
title: "QC Report"
date: "Updated: `r format(Sys.time())`"
geometry: margin=2cm
output:
      pdf_document:
        highlight: tango
        # number_sections: yes
        # toc: yes
        # toc_depth: 2
        # toc_float: yes
params:
  experiment: NA
  run_info: NA
  season: NA
  amp.match.summary: NA
  sum_matched_df: NA
  results: NA
  classification: NA
  read_quality: NA
  cycl_qual_plot: NA
  seq_cont_per_cycle: NA
  read_freq_plot: NA
  # results.summary: NA
  seq.metrics: NA
  base_calls_plot: NA
  ind_na: NA
#header-includes:
#  - \usepackage{pdflscape}
#  - \usepackage{longtable}
---

## `r paste("Experiment:", params$experiment)`


```{r include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)
library(RColorBrewer)
library(GGally)
library(scales)
library(viridis)
library(kableExtra)

other_seasons <- c("winter", "spring", "summer", "fall")
other_seasons <- other_seasons[which(!other_seasons %in% season)]

```

\scriptsize

### Run info
```{r echo=FALSE, message=FALSE, warning=FALSE}
params$run_info %>%
  kableExtra::kable(format = "latex")
  # knitr::kable(df_print = "tibble")
```

### Sequencer metrics
```{r echo=FALSE, message=FALSE, warning=FALSE}
params$seq.metrics %>%
  kableExtra::kable(format = "latex")
  # knitr::kable()
```

### Alignment summary
```{r echo=FALSE, message=FALSE, warning=FALSE}
params$sum_matched_df %>%
  kableExtra::kable(format = "latex")
  # knitr::kable()
```

### Sample Classification
```{r echo=FALSE, message=FALSE, warning=FALSE}
params$classification %>%
  group_by(classification) %>%
  summarise(total = n()) %>%
  kableExtra::kable(format = "latex")
```

\normalsize

```{r echo=FALSE, message=FALSE, warning=FALSE}
##################
# Modify columns #
##################
df <- params$results %>% 
  filter(season == !!params$season)

df$Col=as.factor(gsub('^.', '', df$Sample_Well))
df$Row=factor(gsub('..$', '', df$Sample_Well), levels=rev(toupper(letters[1:8])))
df$Sample=paste0(df$Plate_ID, '-' ,df$Sample_Well)
df$Plate_ID=as.factor(df$Plate_ID)
df$Plate_ID=factor(df$Plate_ID, levels(df$Plate_ID)[order(as.numeric(gsub('Plate', '', levels(df$Plate_ID))))])

###############
# All results #
###############
df_all <- params$results

df_all$Col=as.factor(gsub('^.', '', df_all$Sample_Well))
df_all$Row=factor(gsub('..$', '', df_all$Sample_Well), levels=rev(toupper(letters[1:8])))
df_all$Sample=paste0(df_all$Plate_ID, '-' ,df_all$Sample_Well)
df_all$Plate_ID=as.factor(df_all$Plate_ID)
df_all$Plate_ID=factor(df_all$Plate_ID, levels(df_all$Plate_ID)[order(as.numeric(gsub('Plate', '', levels(df_all$Plate_ID))))])
```

## Plate Map Plot

```{r echo=FALSE, fig.height=10, fig.width=12, message=FALSE, warning=FALSE}
df %>%
  filter(!is.na(Plate_ID)) %>% 
  ggplot(aes(x=Col, y=Row, fill=log10(Count))) + 
  # geom_raster() +
  geom_tile() +
  coord_equal() +
  facet_grid(amplicon~Plate_ID) +
  scale_fill_viridis_c(option = 'plasma') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5))

```

\newpage

## 384 Index Plot

```{r echo=FALSE, fig.height=10, fig.width=12, message=FALSE, warning=FALSE}
myPalette <- colorRampPalette(brewer.pal(9, "Greys"))

df %>%
  filter(!is.na(pm_384)) %>%
  # filter(!is.na(amplicon)) %>% 
  mutate(Count = log10(Count)) %>% 
  # filter(Count > 1) %>% 
  ggplot(aes(x=col_384, y=row_384, fill = Count)) +
  # geom_raster() +
  geom_tile() +
  coord_equal() +
  facet_grid(amplicon~pm_384) +
  scale_x_continuous(breaks = 1:24) +
  scale_y_continuous(breaks = 1:16, trans = "reverse") +
  # coord_cartesian(xlim = c(1,24), ylim = c(1,16)) +
  # ylim(1,16) +
  # xlim(1,24) +
  # scale_fill_gradientn(colours = myPalette(100)) +
  scale_fill_viridis_c(option = 'plasma', na.value = "white") +
  xlab("384 Plate Column") +
  ylab("384 Plate Row") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5))

```

\newpage

## Sample Categorization
Classifies samples as COVID positive/negative, or failure type. Table at end of document.

```{r echo=FALSE, fig.height=10, fig.width=12, message=FALSE, warning=FALSE}
cat_cols <- c("COVID_pos" = "#E69F00", 
              "COVID_neg" = "#0072B2", 
              "failed: low RPP30" = "#009E73", 
              "failed: low S2" = "#000000", 
              "failed: low S2 & RPP30" = "#999999")

classification <- params$classification %>%
  # dplyr::select(index, index2, pm_384, row_384, col_384, Plate_ID, Sample_Well, S2_spike, S2, RPP30, s2_vs_spike, classification) %>% 
  filter(season == !!params$season) %>% 
  mutate(Col = as.factor(gsub('^.', '', Sample_Well)),
         Row = factor(gsub('..$', '', Sample_Well), levels=rev(toupper(letters[1:8]))),
         Plate_ID = factor(Plate_ID, levels = unique(Plate_ID)[order(as.numeric(gsub('Plate', '', unique(Plate_ID))))]))

classification %>% 
  ggplot(aes(x=Col, y=Row, fill= classification)) + 
  geom_tile() +
  coord_equal() +
  # geom_point(data = filter(classification, Sample_Well %in% c("A01","B01")),
  #            aes(color = RPP30 < 10,
  #                shape = RPP30 < 10),
  #            stroke = 1,
  #            show.legend = FALSE) +
  geom_point(aes(color = ctrl_wells, shape = ctrl_wells), 
             stroke = 1,
             show.legend = FALSE) +
  scale_shape_manual(values = c(4, 3)) +
  scale_color_manual(values = c("red","green")) +
  scale_fill_manual(values = cat_cols) +
  facet_wrap(~Plate_ID, ncol = 4) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5),
        legend.position = "top")
```

\newpage

## Read Counts per Well

```{r echo=FALSE, fig.height=10, fig.width=12, message=FALSE, warning=FALSE}
df %>%
  filter(!is.na(Plate_ID)) %>% 
  filter(!is.na(amplicon)) %>% 
  group_by(Plate_ID, Col, Row) %>% 
  summarise(read_counts = sum(Count)) %>% 
  ggplot(aes(x=Col, y=Row, fill=log10(read_counts))) + 
  geom_tile() +
  coord_equal() +
  facet_wrap(~Plate_ID, ncol = 4) +
  scale_fill_viridis_c(option = 'plasma') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5))
```

\newpage

## Sample Caryover

### `r str_to_title(other_seasons[1])`

```{r echo=FALSE, fig.height=10, fig.width=12, message=FALSE, warning=FALSE}
myPalette <- colorRampPalette(brewer.pal(9, "Greys"))

df_all %>%
  filter(!is.na(pm_384),
         season == other_seasons[1]) %>%
  mutate(Count = log10(Count)) %>% 
  ggplot(aes(x=col_384, y=row_384, fill = Count)) +
  geom_tile() +
  coord_equal() +
  facet_grid(amplicon~pm_384) +
  scale_x_continuous(breaks = 1:24) +
  scale_y_continuous(breaks = 1:16, trans = "reverse") +
  scale_fill_viridis_c(option = 'plasma', na.value = "white") +
  xlab("384 Plate Column") +
  ylab("384 Plate Row") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5))

```

\newpage

### `r str_to_title(other_seasons[2])`

```{r echo=FALSE, fig.height=10, fig.width=12, message=FALSE, warning=FALSE}
myPalette <- colorRampPalette(brewer.pal(9, "Greys"))

df_all %>%
  filter(!is.na(pm_384),
         season == other_seasons[2]) %>%
  mutate(Count = log10(Count)) %>% 
  ggplot(aes(x=col_384, y=row_384, fill = Count)) +
  geom_tile() +
  coord_equal() +
  facet_grid(amplicon~pm_384) +
  scale_x_continuous(breaks = 1:24) +
  scale_y_continuous(breaks = 1:16, trans = "reverse") +
  scale_fill_viridis_c(option = 'plasma', na.value = "white") +
  xlab("384 Plate Column") +
  ylab("384 Plate Row") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5))

```

\newpage

### `r str_to_title(other_seasons[3])`

```{r echo=FALSE, fig.height=10, fig.width=12, message=FALSE, warning=FALSE}
myPalette <- colorRampPalette(brewer.pal(9, "Greys"))

df_all %>%
  filter(!is.na(pm_384),
         season == other_seasons[3]) %>%
  mutate(Count = log10(Count)) %>% 
  ggplot(aes(x=col_384, y=row_384, fill = Count)) +
  geom_tile() +
  coord_equal() +
  facet_grid(amplicon~pm_384) +
  scale_x_continuous(breaks = 1:24) +
  scale_y_continuous(breaks = 1:16, trans = "reverse") +
  scale_fill_viridis_c(option = 'plasma', na.value = "white") +
  xlab("384 Plate Column") +
  ylab("384 Plate Row") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=0.5))

```

\newpage

## Classification vs Read Counts per Well

```{r echo=FALSE, fig.height=10, fig.width=12, message=FALSE, warning=FALSE}
df %>%
  filter(!is.na(Plate_ID)) %>% 
  filter(!is.na(amplicon)) %>% 
  group_by(Plate_ID, Col, Row) %>% 
  summarise(read_counts = sum(Count)) %>% 
  left_join(classification, by = c("Plate_ID","Col","Row")) %>% 
  ggplot(aes(x = read_counts, fill = classification)) + 
  geom_histogram(bins = 50) +
  facet_wrap(~classification, ncol = 1, scales = "free_y") +
  scale_x_continuous(trans = pseudo_log_trans(base = 10), breaks = c(0,10,100,1000,1E4,1E5, 1E6, 1E7)) +
  annotation_logticks(sides = "b") +
  theme_bw()
```

\newpage

## Index Swapping

```{r echo=FALSE, message=FALSE, warning=FALSE}
df_all %>% 
  filter(!is.na(index),
         !is.na(index2),
         !is.na(amplicon)) %>% 
  group_by(index, index2) %>% 
  summarise(n = sum(Count)) %>% 
  ggplot(aes(y=index, x=index2, fill=log10(n))) + 
  # geom_raster() +
  geom_tile() +
  scale_fill_gradient(
    low = "#FFFFFF",
    high = "#000000"
  ) +
  # scale_fill_viridis_c(option = 'plasma') +
  coord_equal() +
  # facet_wrap(~amplicon) +
  theme_classic() +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

\newpage

## Swap events for total counts

```{r echo=FALSE, message=FALSE, warning=FALSE}
#scatter plots of s2 reads vs swaps where swaps are the marginal counts (or rates) of swapped reads for each of the indices shared for each sample
mat_df <- df_all %>% 
  group_by(index, index2) %>%
  # filter(amplicon == "RPP30") %>%
  summarise(Count = sum(Count, na.rm = TRUE)) %>%
  ungroup() %>% 
  dplyr::select(index, index2, Count)



missing_ind1 <- tibble(index = levels(df_all$index)[which(!levels(mat_df$index) %in% unique(mat_df$index))],
                       index2 = NA,
                       Count = 0)
missing_ind2 <- tibble(index = NA,
                       index2 = levels(df_all$index2)[which(!levels(mat_df$index2) %in% unique(mat_df$index2))],
                       Count = 0)

mat_df <- rbind(mat_df, missing_ind1, missing_ind2) %>% 
  pivot_wider(values_from = Count, names_from = index)
row_names <- mat_df$index2
mat_df <- as.matrix(mat_df[,2:ncol(mat_df)])
rownames(mat_df) <- row_names

mat_df[which(is.na(mat_df))] <- 0

col_order <- levels(df_all$index)[levels(df_all$index) %in% colnames(mat_df)]
row_order <- levels(df_all$index2)[levels(df_all$index2) %in% rownames(mat_df)]

mat_df <- mat_df[row_order,col_order]

# pheatmap::pheatmap(mat_df, cluster_rows = F, cluster_cols = F, show_colnames = F, show_rownames = F)

UDIs <- diag(mat_df)
diag(mat_df) <- 0
ind1_mismatch <- apply(mat_df, 2, function(x) sum(x))
ind2_mismatch <- apply(mat_df, 1, function(x) sum(x))

swap_events <- rbind(
  tibble(
    UDIs = UDIs,
    index_mismatch = ind1_mismatch,
    index = "index1"
  ),
  tibble(
    UDIs = UDIs,
    index_mismatch = ind2_mismatch,
    index = "index2"
  )
) 

swap_events %>% 
  ggplot(aes(x = UDIs, y = index_mismatch)) +
  geom_point() +
  scale_x_continuous(trans = pseudo_log_trans(base = 10), breaks = c(0,10,100,1000,1E4,1E5, 1E6, 1E7, 1E8)) +
  scale_y_continuous(trans = pseudo_log_trans(base = 10), breaks = c(0,10,100,1000,1E4,1E5, 1E6, 1E7, 1E8)) +
  # coord_fixed(ratio = 1, xlim = NULL, ylim = NULL, expand = TRUE, clip = "on") +
  facet_wrap(~index, ncol = 1) +
  theme_minimal()
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
swap_events %>% 
  mutate(udis_vs_mismatch = UDIs / index_mismatch) %>% 
  ggplot(aes(x = udis_vs_mismatch)) +
  geom_histogram(bins = 100) +
  scale_x_log10() +
  facet_wrap(.~index, ncol = 1) +
  xlab("UDIs / Swap Counts") +
  theme_minimal()
```

\newpage

## RPP30 / S2 counts
Red line = log2(0.5)

```{r echo=FALSE, message=FALSE, warning=FALSE}
params$classification %>% 
  mutate(rpp30_vs_s2 = (RPP30 + 1) / (S2 + S2_spike + 1)) %>% 
  ggplot(aes(x = log2(rpp30_vs_s2))) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = -1, color = "red", linetype = "dashed") +
  xlab("log2 (RPP30 + 1) / (S2 + S2 Spike + 1)") +
  # facet_wrap(~exp, ncol = 1) +
  theme_minimal()
```

\newpage 

## Cycle-specific Quality Distribution

```{r echo=FALSE, message=FALSE}
eval(params$cycl_qual_plot + 
  theme_minimal() + 
  scale_fill_viridis_c(limits = c(0,40)))
```

## Sequence quality per base/cycle

```{r echo=FALSE}
eval(params$read_quality)
```

## Base calls per cycle

```{r echo=FALSE}
eval(params$base_calls_plot)
```

## Read frequency plot

```{r echo=FALSE}
eval(params$read_freq_plot)
```

\newpage

\tiny

```{r echo=FALSE, message=FALSE, warning=FALSE}
#classification %>%
#  filter(classification == "COVID_pos") %>%
#  dplyr::select(-index, -index2, -Plate_384_Number, -Col, -Row) %>% 
#  dplyr::rename(quad = Plate_384_Quadrant) %>% 
#  arrange(Plate_ID, Sample_Well) %>% 
#  # mutate(
#  # RPP30 = cell_spec(RPP30, color = ifelse(RPP30 < 10, "white", "black"),
#  #                 background = ifelse(RPP30 < 10, "red", "white"))
#  # ) %>%
#  # kable(escape = F, booktabs = F) #%>%
#  # kable_styling()
#  kableExtra::kable(format = "latex", longtable = TRUE, caption = "COVID Positive Samples") %>% 
#  kable_styling(latex_options = c("hold_position", "repeat_header")) %>% 
#  landscape()
```

\newpage

```{r echo=FALSE, message=FALSE, warning=FALSE}
#classification %>%
#  filter(grepl("failed:", classification)) %>%
#  dplyr::select(-index, -index2, -s2_vs_spike, -Plate_384_Number, -Col, -Row) %>%
#  dplyr::rename(quad = Plate_384_Quadrant) %>% 
#  arrange(Plate_ID, Sample_Well) %>% 
#  # mutate(
#  # RPP30 = cell_spec(RPP30, color = ifelse(RPP30 < 10, "white", "black"),
#  #                 background = ifelse(RPP30 < 10, "red", "white"))
#  # ) %>%
#  # kable(escape = F, booktabs = F) #%>%
#  # kable_styling()
#  kableExtra::kable(format = "latex", longtable = TRUE, caption = "Inconclusive Samples") %>% 
#  kable_styling(latex_options = c("hold_position", "repeat_header")) %>% 
#  # column_spec(1:16,width = "1in") %>%
#  landscape()
```

\newpage

```{r echo=FALSE, message=FALSE, warning=FALSE}
#classification %>%
#  filter(classification == "COVID_neg") %>%
#  dplyr::select(-index, -index2, -Plate_384_Number, -Col, -Row) %>% 
#  dplyr::rename(quad = Plate_384_Quadrant) %>% 
#  arrange(Plate_ID, Sample_Well) %>% 
#  # mutate(
#  # RPP30 = cell_spec(RPP30, color = ifelse(RPP30 < 10, "white", "black"),
#  #                 background = ifelse(RPP30 < 10, "red", "white"))
#  # ) %>%
#  # kable(escape = F, booktabs = F) #%>%
#  # kable_styling()
#  kableExtra::kable(format = "latex", longtable = TRUE, caption = "COVID Negative Samples") %>% 
#  kable_styling(latex_options = c("hold_position", "repeat_header")) %>% 
#  landscape()
```

\normalsize

```{r echo=FALSE, message=FALSE, warning=FALSE}
sessionInfo()
```





